Pretty darn good control: when are approximate solutions better than approximate models

### 2023-01-20 manuscript/manuscript.Rmd

authors: 
- name: Felipe Montealegre-Mora
  address: Department of Environmental Science, Policy, and Management, University of California, Berkeley
- name: Marcus Laperolerie
  address: Department of Environmental Science, Policy, and Management, University of California, Berkeley
- name: Melissa Chapman
  address: Department of Environmental Science, Policy, and Management, University of California, Berkeley
- name: Abigail G. Keller
  address: Department of Environmental Science, Policy, and Management, University of California, Berkeley
- name: Carl Boettiger
  address: Department of Environmental Science, Policy, and Management, University of California, Berkeley
  email: cboettig@berkele.edu

@@

Despite the complexity inherent in natural ecosystems, analytically tractable models have long played an essential role in their understanding and management.
Ecological management may be understood as an optimal control problem - a manager must determine a sequence of actions (a "policy") to optimize some desired objective(s) and constraints (e.g. conservation goals, natural resource extraction, economic and social constraints.)

@@

Our results suggest that DRL can find solutions that are more sustainable *and* more profitable than both MSY and TAC.
Indeed, DRL finds solutions in which, for certain periods of time, no harvest is performed on the system.
These periods effectively allow the system to, in a sense, ‘‘reset’’ itself.
They lead to a distinctive oscillating behavior of the system state over time, with periods of relative depletion followed by periods of growth. 
Figure \ref{fig:tbd} exemplifies this point by comparing one times series of the system state obtained using each of these policies.

@@

The title of this piece references a mathematical biology workshop at NIMBioS
organized by Paul Armsworth, Alan Hastings, Megan Donahue, and Carl Towes in
2011 that first posed the question addressed here.

>>>>

authors: 
- name: Felipe Montealegre-Mora
  address: Department of Environmental Science, Policy, and Management, University of California, Berkeley
- name: Marcus Laperolerie
  address: Department of Environmental Science, Policy, and Management, University of California, Berkeley
- name: Melissa Chapman
  address: Department of Environmental Science, Policy, and Management, University of California, Berkeley
- name: Abigail G. Keller
  address: Department of Environmental Science, Policy, and Management, University of California, Berkeley
- name: Carl Boettiger
  address: Department of Environmental Science, Policy, and Management, University of California, Berkeley
  email: cboettig@berkeley.edu

@@

Ecological management may be understood as an optimal control problem - a manager must determine a sequence of actions (a "policy") to optimize some desired objective(s) and constraints (e.g. conservation goals, natural resource extraction, economic and social constraints.)
Despite the complexity inherent in natural ecosystems, analytically tractable models have long played an essential role in their understanding and management.
...

<!-- Why fisheries-focused examples -->
Global fisheries management has long been a crucible for both the development of theory and best practices in actual management for many reasons. Fisheries are an essential natural resource that provide the primary source of protein for one in every four humans, and have faced widely documented declines due to overfishing [@Worm2006, @Costello2016]. 
Because of this importance, fisheries also represent some of the longest and best studied timeseries of population dynamics available to ecologists, as well as a correspondingly rich area for the development of theory and practice.
Methods developed in the study and management of fisheries are frequently applied to other areas of ecological management, from concepts such as maximum sustainable yield [], optimal control under observation uncertainty [@Memarzadeh2019], to adaptive management [@Walters], management strategy evaluation [@Punt], to the studies of conservation policy and international negotiations [@].

<!-- A sweeping summary of fisheries decision theory, don't bite me-->
To understand the importance of a potential application of novel data-driven methods to the real world, it is most instructive to have a clearer picture of how models are currently used to inform ecological management in fisheries.
In any issue as large and contentious as fisheries management, it may be far more difficult to provide a generalization  about how things actually work than to make sweeping recommendations about how it ought to.
All the same, it is our firm belief that any researcher seeking to do the latter should first attempt to understand the former, so we beg the forbearance of any specialists in the practice as we seek to provide a concise overview of fisheries management today.
Here, we focus specifically on practices in data rich, 'well-managed' fisheries, which, despite some issues, may be among the best examples of ecological management at continental scale that is largely driven by scientific processes, models and data.

<!-- Fisheries population models are rich but decisions aren't, mostly constant mortality -->
While fisheries stock assessment models are quite rich, the choice of possible strategies remains motivated by intution from much simpler, analytically tractable one-dimensional models.
This fact is not always obvious at cursory inspection, and can be met with cynicism even among professional fisheries managers -- after all, typical stock assessment models have over 100 parameters (e.g. see the stock assessments referenced from in @RAMLegacy) and best practices such as management strategy evaluation (MSE) evaluate and select potential policies by comparing across large and diverse simulations [e.g. @Punt2016].
However, the space of policies considered under such approaches is necessarily far smaller than the set of all possible action sequences.
Instead, fisheries management approaches such as MSE rely on theory based on more tractable models to restrict the search space of possible policies.  
Specifically, most fisheries policies are framed in terms of a *constant mortality* policy, (sometimes a simple piece-wise linear function, i.e. a single linear segment, is introduced as a precaution a very low stock sizes), where fishing 'mortality' is defined as the ratio of harvest (catch) to stock size, $F = H_t/x_t$.

The intuition for a constant mortality strategy is simple and well established

Constant mortality policies have a strong basis in theory as well as practical justifications.

<!-- In contrast to constant mortality, the possible decision space is huge-->
The management of any ecological system can be framed in terms of a Sequential Decision Process - 
the manager observes the ecosystem state ($x_t$) and decides on action ($a_t$) to take (including potentially no action), potentially receiving any costs or benefits associated with that action and and state (reward $r_t$  $r_t = U(x_t,a_t)$),
while the system evolves into a new state in response $x_{t+1} = f(x_t, a_t)$. When the probability of the future state can be determined by the current state and action, we say the system is Markovian and the process a Markov Decision Process (MDP).
As in a game of chess, the space of possible sequences of actions a manager might choose over any realization of the Markov process is enormous.
When the action and state space are discrete and sufficiently small, computational approaches such as dynamic programming may be used to guarantee an exhaustive search for a single optimal policy.
Dynamic programming scales very poorly to larger state and action spaces.

Deep reinforcement learning has opened up increasingly compelling alternative

Effort $Q = E X_t$

Fisheries policies are typically implemented in terms of *catch quotas*, i.e. a limit to the number of metric tonnes of fish that can be caught during a given season.

and/or *effort controls*, restrictions on the fishing effort -- in numbers of boats, time of year, or days at sea

@@

Our results suggest that DRL can find solutions that are more sustainable *and* more profitable than both MSY and TAC.
Indeed, DRL finds solutions in which, for certain periods of time, no harvest is performed on the system.
These periods effectively allow the system to, in a sense, "reset" itself.
They lead to a distinctive oscillating behavior of the system state over time, with periods of relative depletion followed by periods of growth. 
Figure \ref{fig:tbd} exemplifies this point by comparing one times series of the system state obtained using each of these policies.

@@

The title of this piece references a mathematical biology workshop at NIMBioS
organized by Paul Armsworth, Alan Hastings, Megan Donahue, and Carl Towes in
2011 which first sought to emphasize 'pretty darn good' control solutions to
more realistic problems over optimal control to idealized ones. 
This material is based upon work supported by the National Science Foundation under Grant No. DBI-1942280.
